---
title: "Surveys:Rarefaction_2025"
output:
  html_document:
    df_print: paged
date: "2026-03-05"
---

## Load packages
```{r}
setwd("~/Library/CloudStorage/Box-Box/DesktopB/Viscaria Data/Viscaria_2025")

library(tidyverse)
library(readxl)
library(readr)
library(here)
library(reshape2)
library(janitor) 
library(vegan)
library(ggplot2)
library(RColorBrewer)
library(nlme)
library(pals)
library(imputeTS)
library(ggforce)
library(scales)
library(randomcoloR)
library(vegan)
library(fossil)
```

# *Since 2021, how many species did we see as we increased observation bouts regardless of population?*

###Load and clean data (all years clubbed, modified observation IDs)
```{r load-prep-allpop, message=FALSE, warning=FALSE}
##SPECIES
#Load data (all years clubbed, modified observation IDs)
sac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
sac_allpop <- sac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Observation_ID_year = paste(Year, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Observation_ID_year, .after = Observation_ID)

#Remove populations - club them all and only look at observation ID_year combo
sac_allpop_counts <- sac_allpop %>% group_by(Observation_ID_year, Pollinator_Genus, Pollinator_Species) %>%
  summarise(Total_Individuals = sum(Individuals, na.rm = T), .groups = "drop_last")%>%
  unite(Pollinator_ID, c(Pollinator_Genus, Pollinator_Species), remove = F, na.rm = T) %>%
  mutate(Pollinator_ID = ifelse(Pollinator_ID == "", "Unknown", Pollinator_ID))

#Restructure data to wide format
restructured_dataset <-pivot_wider(sac_allpop_counts, 
              id_cols = Observation_ID_year,
              names_from = Pollinator_ID, 
              values_from = Total_Individuals)
restructured_dataset[is.na(restructured_dataset)] <- 0

sac_allpop_counts_res <- restructured_dataset %>%
  ungroup() %>%
  select(-Observation_ID_year)

#---- 
#FUNCTIONAL GROUPS
#Load data (all years clubbed, modified observation IDs)
fgac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
fgac_allpop <- fgac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Observation_ID_year = paste(Year, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Observation_ID_year, .after = Observation_ID)

#Remove populations - club them all and only look at observation ID_year combo
fgac_allpop_counts <- fgac_allpop %>% group_by(Observation_ID_year, Pollinator_Group_Collapsed) %>%
  summarise(Total_Individuals = sum(Individuals, na.rm = T), .groups = "drop_last")

#Restructure data to wide format
restructured_dataset <-pivot_wider(fgac_allpop_counts, 
              id_cols = Observation_ID_year,
              names_from = Pollinator_Group_Collapsed, 
              values_from = Total_Individuals)
restructured_dataset[is.na(restructured_dataset)] <- 0

fgac_allpop_counts_res <- restructured_dataset %>%
  ungroup() %>%
  select(-Observation_ID_year)
```

###Calculating Richeness and SD (and plotting clubbed Species Accumulation Curves) 
```{r sac-allpop, message=FALSE, warning=FALSE}
#SPECIES
# Fit SAC: if we randomly add one more bout, how many species have we accumulated?
sac_allpop_rarefaction <- specaccum(sac_allpop_counts_res, method="random",permutations=100) 

# convert to df compatible with ggplot
rarefaction_df <- data.frame(bouts = sac_allpop_rarefaction$sites, 
                             richness = sac_allpop_rarefaction$richness, 
                             sd = sac_allpop_rarefaction$sd)

#to make compatible with sep year df
sac_allpop_allyears <- rarefaction_df %>%
  mutate(Year = "All Years") %>%
  relocate(Year, .before = bouts)

#plotting here is redundant, will plot with all sep years later
#----
## BASE R
#plot mean richness and confidence intervals as we do more bouts = species accumulation curve
plot(sac_allpop_rarefaction, ci.type="line", col="black", lwd=2, ci.lty=2, ci.col="gray80",
ylab="Species (N)", xlab = "Observation Bouts", main = "Rarefaction Curve 2021-2025")

# and the distribution of richness in every simulation as we do more bouts
#At observation bout X, we found between Y and Z species depending on the order we encountered sites — the boxplot shows that uncertainty.
boxplot(sac_allpop_rarefaction, col="yellow", pch="", ann = FALSE)
title(ylab = "Species (N)", xlab = "Observation Bouts")

## GGPLOT 
#simple line with confidence interval
ggplot(rarefaction_df, aes(x = bouts, y = richness)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd),
              fill = "gray80", alpha = 0.5) +
  geom_line(color = "black", linewidth = 1) +
  labs(x = "Observation Bouts",
    y = "Species (N)",
    title = "Rarefaction Curve") +
  theme_classic()
# with boxplots complicated, just use base R for now

#----
#FUNCTIONAL GROUPS
# Fit SAC: if we randomly add one more bout, how many species have we accumulated?
fgac_allpop_rarefaction <- specaccum(fgac_allpop_counts_res, method="random",permutations=100) 

# convert to df compatible with ggplot
rarefaction_df <- data.frame(bouts = fgac_allpop_rarefaction$sites, 
                             richness = fgac_allpop_rarefaction$richness, 
                             sd = fgac_allpop_rarefaction$sd)
#to make compatible with sep year df
fgac_allpop_allyears <- rarefaction_df %>%
  mutate(Year = "All Years") %>%
  relocate(Year, .before = bouts)

#plotting here is redundant, will plot with all sep years later
#----
## BASE R
#plot mean richness and confidence intervals as we do more bouts = species accumulation curve
plot(fgac_allpop_rarefaction, ci.type="line", col="black", lwd=2, ci.lty=2, ci.col="gray80",
ylab="Functional Groups (N)", xlab = "Observation Bouts", main = "Rarefaction Curve 2021-2025")

# and the distribution of richness in every simulation as we do more bouts
#At observation bout X, we found between Y and Z species depending on the order we encountered sites — the boxplot shows that uncertainty.
boxplot(fgac_allpop_rarefaction, col="yellow", pch="", ann = FALSE)
title(ylab = "Functional Groups (N)", xlab = "Observation Bouts")

## GGPLOT 
#simple line with confidence interval
ggplot(rarefaction_df, aes(x = bouts, y = richness)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd),
              fill = "gray80", alpha = 0.5) +
  geom_line(color = "black", linewidth = 1) +
  labs(x = "Observation Bouts",
    y = "Functional Groups (N)",
    title = "Rarefaction Curve") +
  theme_classic()
# with boxplots complicated, just use base R for now
```

## In each year, how many species did we see as we increased observation bouts regardless of population?
###Load data and clean (all years separated, modified observation IDs)
```{r load-prep-allpop-byyear, message=FALSE, warning=FALSE}
#Load data (all years clubbed, modified observation IDs)
sac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
sac_allpop <- sac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Observation_ID_year = paste(Year, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Observation_ID_year, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each year and bind into single dataframe
years <- c(2021, 2022, 2023, 2024, 2025)
sac_allpop_byyear <- lapply(years, function(yr) {
  
  df <- sac_allpop %>%
    filter(Year == yr) %>%
    group_by(Observation_ID_year, Pollinator_Genus, Pollinator_Species) %>%
    summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
    unite(Pollinator_ID, c(Pollinator_Genus, Pollinator_Species), remove = FALSE, na.rm = TRUE) %>%
    mutate(Pollinator_ID = ifelse(Pollinator_ID == "", "Unknown", Pollinator_ID)) %>%
    pivot_wider(id_cols = Observation_ID_year,
                names_from = Pollinator_ID,
                values_from = Total_Individuals) %>%
    ungroup() %>%
    select(-Observation_ID_year) %>%
    replace(is.na(.), 0)
  
  sac <- specaccum(df, method = "random", permutations = 100)
  
  data.frame(Year = as.factor(yr),
    bouts    = sac$sites,
    richness = sac$richness,
    sd       = sac$sd)}) %>% 
bind_rows()

#bind allyears with sep years
sac_allpop_combo <- bind_rows(sac_allpop_byyear,sac_allpop_allyears) 

#----
#FUNCTIONAL GROUPS
#Load data (all years clubbed, modified observation IDs)
fgac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
fgac_allpop <- fgac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Observation_ID_year = paste(Year, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Observation_ID_year, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each year and bind into single dataframe
years <- c(2021, 2022, 2023, 2024, 2025)
fgac_allpop_byyear <- lapply(years, function(yr) {
  
  df <- fgac_allpop %>%
    filter(Year == yr) %>%
    group_by(Observation_ID_year, Pollinator_Group_Collapsed) %>%
    summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(id_cols = Observation_ID_year,
                names_from = Pollinator_Group_Collapsed,
                values_from = Total_Individuals) %>%
    ungroup() %>%
    select(-Observation_ID_year) %>%
    replace(is.na(.), 0)
  
  sac <- specaccum(df, method = "random", permutations = 100)
  
  data.frame(Year = as.factor(yr),
    bouts    = sac$sites,
    richness = sac$richness,
    sd       = sac$sd)}) %>% 
bind_rows()

#bind allyears with sep years
fgac_allpop_combo <- bind_rows(fgac_allpop_byyear,fgac_allpop_allyears) 
```

### Plotting all Species Accumulation Curves (sep years and clubbed)
```{r sac-allpop-byyear, message=FALSE, warning=FALSE}

# replace with sac_allpop_byyear for sep years
ggplot(sac_allpop_combo, aes(x = bouts, y = richness, color = Year, fill = Year)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  labs(x = "Observation Bouts", y = "Species (N)",
       title = "Rarefaction Curves by Year") +
  theme_classic()
# replace with fgac_allpop_byyear for sep years
ggplot(fgac_allpop_combo, aes(x = bouts, y = richness, color = Year, fill = Year)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  labs(x = "Observation Bouts", y = "Functional Groups (N)",
       title = "Rarefaction Curves by Year") +
  theme_classic()

```


# *Since 2021, how many species did we see as we increased number of populations?*

###Load and clean data (all years and populations clubbed, modified observation IDs)
```{r load-prep-allpop, message=FALSE, warning=FALSE}
##SPECIES
#Load data (all observation bouts clubbed, modified observation IDs)
sac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
sac_allpop <- sac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Year = paste(Population, Year, sep = "_")) %>%
  relocate(Observation_ID_new, Population_Year, .after = Observation_ID)

#Remove observation bouts - club them all within pop and year and only look at pop_year combo
sac_allpop_counts <- sac_allpop %>% group_by(Population_Year, Pollinator_Genus, Pollinator_Species) %>%
  summarise(Total_Individuals = sum(Individuals, na.rm = T), .groups = "drop_last")%>%
  unite(Pollinator_ID, c(Pollinator_Genus, Pollinator_Species), remove = F, na.rm = T) %>%
  mutate(Pollinator_ID = ifelse(Pollinator_ID == "", "Unknown", Pollinator_ID))

#Restructure data to wide format
restructured_dataset <-pivot_wider(sac_allpop_counts, 
              id_cols = Population_Year,
              names_from = Pollinator_ID, 
              values_from = Total_Individuals)
restructured_dataset[is.na(restructured_dataset)] <- 0

sac_allpop_counts_res <- restructured_dataset %>%
  ungroup() %>%
  select(-Population_Year)

#---- 
#FUNCTIONAL GROUPS
#Load data (all years clubbed, modified observation IDs)
fgac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
fgac_allpop <- fgac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Year = paste(Population, Year, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Year, .after = Observation_ID)

#Remove populations - club them all and only look at observation ID_year combo
fgac_allpop_counts <- fgac_allpop %>% group_by(Population_Year, Pollinator_Group_Collapsed) %>%
  summarise(Total_Individuals = sum(Individuals, na.rm = T), .groups = "drop_last")

#Restructure data to wide format
restructured_dataset <-pivot_wider(fgac_allpop_counts, 
              id_cols = Population_Year,
              names_from = Pollinator_Group_Collapsed, 
              values_from = Total_Individuals)
restructured_dataset[is.na(restructured_dataset)] <- 0

fgac_allpop_counts_res <- restructured_dataset %>%
  ungroup() %>%
  select(-Population_Year)
```

###Calculating Richeness and SD (and plotting clubbed Species Accumulation Curves) 
```{r sac-allpop, message=FALSE, warning=FALSE}
#SPECIES
# Fit SAC: if we randomly add one more population, how many species have we accumulated? (assume each population is new every year)
sac_allpop_rarefaction <- specaccum(sac_allpop_counts_res, method="random",permutations=100) 

# convert to df compatible with ggplot
rarefaction_df <- data.frame(pops = sac_allpop_rarefaction$sites, 
                             richness = sac_allpop_rarefaction$richness, 
                             sd = sac_allpop_rarefaction$sd)
#to make compatible with sep year df
sac_allpop_allyears <- rarefaction_df %>%
  mutate(Year = "All Years") %>%
  relocate(Year, .before = pops)

#plotting here is redundant, will plot with all sep years later
#----
##BASE R
#plot mean richness and confidence intervals as we sample more populations = species accumulation curve
plot(sac_allpop_rarefaction, ci.type="line", col="black", lwd=2, ci.lty=2, ci.col="gray80",
ylab="Species (N)", xlab = "Populations", main = "Rarefaction Curve 2021-2025")

# and the distribution of richness in every simulation as we sample more populations
#At population X, we found between Y and Z species depending on the order we encountered sites — the boxplot shows that uncertainty.
boxplot(sac_allpop_rarefaction, col="yellow", pch="", ann = FALSE)
title(ylab = "Species (N)", xlab = "Populations")

## GGPLOT 
#simple line with confidence interval
ggplot(rarefaction_df, aes(x = pops, y = richness)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd),
              fill = "gray80", alpha = 0.5) +
  geom_line(color = "black", linewidth = 1) +
  labs(x = "Populations",
    y = "Species (N)",
    title = "Rarefaction Curve") +
  theme_classic()
# with boxplots complicated, just use base R for now

#----
#FUNCTIONAL GROUPS
# Fit SAC: if we randomly add one more population, how many species have we accumulated?
fgac_allpop_rarefaction <- specaccum(fgac_allpop_counts_res, method="random",permutations=100) 

# convert to df compatible with ggplot
rarefaction_df <- data.frame(pops = fgac_allpop_rarefaction$sites, 
                             richness = fgac_allpop_rarefaction$richness, 
                             sd = fgac_allpop_rarefaction$sd)

#to make compatible with sep year df
fgac_allpop_allyears <- rarefaction_df %>%
  mutate(Year = "All Years") %>%
  relocate(Year, .before = pops)

#plotting here is redundant, will plot with all sep years later
#----
## BASE R
#plot mean richness and confidence intervals as we sample more populations = species accumulation curve
plot(fgac_allpop_rarefaction, ci.type="line", col="black", lwd=2, ci.lty=2, ci.col="gray80",
ylab="Functional Groups (N)", xlab = "Populations", main = "Rarefaction Curve 2021-2025")

# and the distribution of richness in every simulation as we sample more populations
#At population X, we found between Y and Z species depending on the order we encountered sites — the boxplot shows that uncertainty.
boxplot(fgac_allpop_rarefaction, col="yellow", pch="", ann = FALSE)
title(ylab = "Functional Groups (N)", xlab = "Observation Bouts")

## GGPLOT 
#simple line with confidence interval
ggplot(rarefaction_df, aes(x = bouts, y = richness)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd),
              fill = "gray80", alpha = 0.5) +
  geom_line(color = "black", linewidth = 1) +
  labs(x = "Observation Bouts",
    y = "Functional Groups (N)",
    title = "Rarefaction Curve") +
  theme_classic()
# with boxplots complicated, just use base R for now
```



## In each year, how many species did we see as we increased observation bouts regardless of population?
###Load data and clean (all years separated but populations clubbed, modified observation IDs)
```{r load-prep-allpop-byyear, message=FALSE, warning=FALSE}
#Load data (all years clubbed, modified observation IDs)
sac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
sac_allpop <- sac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Year = paste(Population, Year, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Year, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each year and bind into single dataframe
years <- c(2021, 2022, 2023, 2024, 2025)
sac_allpop_byyear <- lapply(years, function(yr) {
  
  df <- sac_allpop %>%
    filter(Year == yr) %>%
    group_by(Population_Year, Pollinator_Genus, Pollinator_Species) %>%
    summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
    unite(Pollinator_ID, c(Pollinator_Genus, Pollinator_Species), remove = FALSE, na.rm = TRUE) %>%
    mutate(Pollinator_ID = ifelse(Pollinator_ID == "", "Unknown", Pollinator_ID)) %>%
    pivot_wider(id_cols = Population_Year,
                names_from = Pollinator_ID,
                values_from = Total_Individuals) %>%
    ungroup() %>%
    select(-Population_Year) %>%
    replace(is.na(.), 0)
  
  sac <- specaccum(df, method = "random", permutations = 100)
  
  data.frame(Year = as.factor(yr),
    pops    = sac$sites,
    richness = sac$richness,
    sd       = sac$sd)}) %>% 
bind_rows()

#bind allyears with sep years
sac_allpop_combo <- bind_rows(sac_allpop_byyear,sac_allpop_allyears) 


#----
#FUNCTIONAL GROUPS
#Load data (all years clubbed, modified observation IDs)
fgac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
fgac_allpop <- fgac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Year = paste(Population, Year, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Year, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each year and bind into single dataframe
years <- c(2021, 2022, 2023, 2024, 2025)
fgac_allpop_byyear <- lapply(years, function(yr) {
  
  df <- fgac_allpop %>%
    filter(Year == yr) %>%
    group_by(Population_Year, Pollinator_Group_Collapsed) %>%
    summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(id_cols = Population_Year,
                names_from = Pollinator_Group_Collapsed,
                values_from = Total_Individuals) %>%
    ungroup() %>%
    select(-Population_Year) %>%
    replace(is.na(.), 0)
  
  sac <- specaccum(df, method = "random", permutations = 100)
  
  data.frame(Year = as.factor(yr),
    pops    = sac$sites,
    richness = sac$richness,
    sd       = sac$sd)}) %>% 
bind_rows()

#bind allyears with sep years
fgac_allpop_combo <- bind_rows(fgac_allpop_byyear,fgac_allpop_allyears) 

```

### Plot all Species Accumulation Curves (sep and clubbed years)
```{r sac-allpop-byyear, message=FALSE, warning=FALSE}
#replace with sac_allpop_byyear for only sep years
ggplot(sac_allpop_combo, aes(x = pops, y = richness, color = Year, fill = Year)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  labs(x = "Populations", y = "Species (N)",
       title = "Rarefaction Curves by Year") +
  theme_classic()

# replace with fgac_allpop_byyear for only sep years
ggplot(fgac_allpop_combo, aes(x = pops, y = richness, color = Year, fill = Year)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  labs(x = "Populations", y = "Functional Groups (N)",
       title = "Rarefaction Curves by Year") +
  theme_classic()

```


# *Since 2021, how many species did we see as we increased observation bouts (and observation minutes/hours), in each population*

###Load and clean data (all years clubbed, modified observation IDs)
```{r load-prep-bypop, message=FALSE, warning=FALSE}
##SPECIES
#Load data (all years separate, modified observation IDs)
sac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
sac_allpop <- sac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Bout = paste(Population, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Bout, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each pop and bind into single dataframe
y#ears <- c(2021, 2022, 2023, 2024, 2025)
populations <- c("Knivsasen", "Hardeberga", "Veberods_Ljung", "Kullamolla", "Hallamolla",
                 "Brosarps_Norra_Backar", "Rorum", "Baskemolla", "Petersborg", "Kvesarum",
                 "Sodra_Rorum", "Bollerup", "Vagasked", "Klingstorp", "Tjurko",
                 "Vasternas", "Hasslo", "Aspo_Madviksvagen", "Sturko", "Gosslunda",
                 "Kleva_Strandskog", "Ynde", "Avaskar", "Hovmantorp", "Uras",
                 "Torne", "Brosarps_Backar", "Animskog_Cliff", "Animskog_Creek", "Aspo_Krokvagen",
                 "Billebjer", "Bokhultet", "Faringtofta", "Hjortahammar", "Jarnavik",
                 "Lyckas", "Mocklo", "Mollerod", "Sjuhalla", "Skaraskog",
                 "Svartbacksmala", "Taxas", "Varaskruv")

sac_bypop <- expand_grid(Population = populations) %>%
  pmap_dfr(function(Population) { #alternate to lapply
    
    # Filter for specific population combo
    df_filtered <- sac_allpop %>%
      filter(Population == !!Population)
    
    # Skip if too few observations to run specaccum
    if (nrow(df_filtered) < 2) return(NULL)
    
    mat <- df_filtered %>%
      group_by(Population_Bout, Pollinator_Genus, Pollinator_Species) %>%
      summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
      unite(Pollinator_ID, c(Pollinator_Genus, Pollinator_Species), remove = FALSE, na.rm = TRUE) %>%
      mutate(Pollinator_ID = ifelse(Pollinator_ID == "", "Unknown", Pollinator_ID)) %>%
      pivot_wider(id_cols = Population_Bout,
                  names_from = Pollinator_ID,
                  values_from = Total_Individuals) %>%
      ungroup() %>%
      select(-Population_Bout) %>%
      replace(is.na(.), 0)
    
    # Skip if matrix has too few rows
    if (nrow(mat) < 2) return(NULL)
    
    sac <- specaccum(mat, method = "random", permutations = 100)
    
    data.frame(
      Population = Population,
      bouts      = sac$sites,
      richness   = sac$richness,
      sd         = sac$sd
    )
  })


#---- 
#FUNCTIONAL GROUPS
#Load data (all years clubbed, modified observation IDs)
fgac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
fgac_allpop <- fgac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Bout = paste(Population, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Bout, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each pop and bind into single dataframe
#years <- c(2021, 2022, 2023, 2024, 2025)
populations <- c("Knivsasen", "Hardeberga", "Veberods_Ljung", "Kullamolla", "Hallamolla",
                 "Brosarps_Norra_Backar", "Rorum", "Baskemolla", "Petersborg", "Kvesarum",
                 "Sodra_Rorum", "Bollerup", "Vagasked", "Klingstorp", "Tjurko",
                 "Vasternas", "Hasslo", "Aspo_Madviksvagen", "Sturko", "Gosslunda",
                 "Kleva_Strandskog", "Ynde", "Avaskar", "Hovmantorp", "Uras",
                 "Torne", "Brosarps_Backar", "Animskog_Cliff", "Animskog_Creek", "Aspo_Krokvagen",
                 "Billebjer", "Bokhultet", "Faringtofta", "Hjortahammar", "Jarnavik",
                 "Lyckas", "Mocklo", "Mollerod", "Sjuhalla", "Skaraskog",
                 "Svartbacksmala", "Taxas", "Varaskruv")

fgac_bypop <- expand_grid( Population = populations) %>%
  pmap_dfr(function(Population) { #alternate to lapply
    
    # Filter for specific population 
    df_filtered <- fgac_allpop %>%
      filter(Population == !!Population)
    
    # Skip if too few observations to run specaccum
    if (nrow(df_filtered) < 2) return(NULL)
    
    mat <- df_filtered %>%
      group_by(Population_Bout, Pollinator_Group_Collapsed) %>%
      summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
      pivot_wider(id_cols = Population_Bout,
                  names_from = Pollinator_Group_Collapsed,
                  values_from = Total_Individuals) %>%
      ungroup() %>%
      select(-Population_Bout) %>%
      replace(is.na(.), 0)
    
    # Skip if matrix has too few rows
    if (nrow(mat) < 2) return(NULL)
    
    fgac <- specaccum(mat, method = "random", permutations = 100)
    
    data.frame(
      Population = Population,
      bouts      = fgac$sites,
      richness   = fgac$richness,
      sd         = fgac$sd
    )
  })

```


### Plotting all Species Accumulation Curves (sep years)
```{r sac-bypop, message=FALSE, warning=FALSE}

#ggplot(sac_bypop, aes(x = bouts, y = richness, color = Population, fill = Population)) +
#convert to hours for more meaningful plots
  ggplot(sac_bypop, aes(x = (bouts*10)/60, y = richness, color = Population, fill = Population)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  #facet_wrap(~ Year, scales = "free_x") +  # free_x because years may have different n bouts
  labs(x = "Observation Hours", y = "Species (N)",
  #labs(x = "Observation Bouts", y = "Species (N)",
       title = "Rarefaction Curves by Population") +
  theme_classic()+
  theme(legend.position = "right",
        legend.text = element_text(size = 5))

#ggsave("rarefaction_bypop_sp.png", width = 20, height = 10, dpi = 300)

#FUNCTIONAL GROUPS
#ggplot(fgac_bypop, aes(x = bouts, y = richness, color = Population, fill = Population)) +
#convert to hours for more meaningful plots
  ggplot(fgac_bypop, aes(x = (bouts*10)/60, y = richness, color = Population, fill = Population)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  #facet_wrap(~ Year, scales = "free_x") +  # free_x because years may have different n bouts
#  labs(x = "Observation Bouts", y = "Functional Groups (N)",
  labs(x = "Observation Hours", y = "Functional Groups (N)",
       title = "Rarefaction Curves by Population") +
  theme_classic()
#ggsave("rarefaction_bypop_fg.png", width = 20, height = 10, dpi = 300)
```

## In each year, how many species did we see as we increased observation bouts, in each population?
###Load data and clean (all years separated, modified observation IDs)
```{r load-prep-bypop-byyear, message=FALSE, warning=FALSE}
#Load data (all years separate, modified observation IDs)
sac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
sac_allpop <- sac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Bout = paste(Population, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Bout, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each year and bind into single dataframe
years <- c(2021, 2022, 2023, 2024, 2025)
populations <- c("Knivsasen", "Hardeberga", "Veberods_Ljung", "Kullamolla", "Hallamolla",
                 "Brosarps_Norra_Backar", "Rorum", "Baskemolla", "Petersborg", "Kvesarum",
                 "Sodra_Rorum", "Bollerup", "Vagasked", "Klingstorp", "Tjurko",
                 "Vasternas", "Hasslo", "Aspo_Madviksvagen", "Sturko", "Gosslunda",
                 "Kleva_Strandskog", "Ynde", "Avaskar", "Hovmantorp", "Uras",
                 "Torne", "Brosarps_Backar", "Animskog_Cliff", "Animskog_Creek", "Aspo_Krokvagen",
                 "Billebjer", "Bokhultet", "Faringtofta", "Hjortahammar", "Jarnavik",
                 "Lyckas", "Mocklo", "Mollerod", "Sjuhalla", "Skaraskog",
                 "Svartbacksmala", "Taxas", "Varaskruv")

sac_bypop_byyear <- expand_grid(Year = years, Population = populations) %>%
  pmap_dfr(function(Year, Population) { #alternate to lapply
    
    # Filter for specific year-population combo
    df_filtered <- sac_allpop %>%
      filter(Year == !!Year, Population == !!Population)
    
    # Skip if too few observations to run specaccum
    if (nrow(df_filtered) < 2) return(NULL)
    
    mat <- df_filtered %>%
      group_by(Population_Bout, Pollinator_Genus, Pollinator_Species) %>%
      summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
      unite(Pollinator_ID, c(Pollinator_Genus, Pollinator_Species), remove = FALSE, na.rm = TRUE) %>%
      mutate(Pollinator_ID = ifelse(Pollinator_ID == "", "Unknown", Pollinator_ID)) %>%
      pivot_wider(id_cols = Population_Bout,
                  names_from = Pollinator_ID,
                  values_from = Total_Individuals) %>%
      ungroup() %>%
      select(-Population_Bout) %>%
      replace(is.na(.), 0)
    
    # Skip if matrix has too few rows
    if (nrow(mat) < 2) return(NULL)
    
    sac <- specaccum(mat, method = "random", permutations = 100)
    
    data.frame(
      Population = Population,
      Year       = as.factor(Year),
      bouts      = sac$sites,
      richness   = sac$richness,
      sd         = sac$sd
    )
  })

#----
#FUNCTIONAL GROUPS
#Load data (all years clubbed, modified observation IDs)
fgac_allpop <- read_excel(here("Data_Clean", "Surveys", "Pollinator_Observations_2125_Clean.xlsx"),
                         na= c("NA", ""))
fgac_allpop <- fgac_allpop %>%
  group_by(Year) %>%
  mutate(Observation_ID_new = as.integer(factor(Observation_ID))) %>%
  ungroup() %>%
  mutate(Population_Bout = paste(Population, Observation_ID_new, sep = "_"))%>%
  relocate(Observation_ID_new, Population_Bout, .after = Observation_ID)


# Use loop to clean, restracture widely and run specaccum for each year and bind into single dataframe
years <- c(2021, 2022, 2023, 2024, 2025)
populations <- c("Knivsasen", "Hardeberga", "Veberods_Ljung", "Kullamolla", "Hallamolla",
                 "Brosarps_Norra_Backar", "Rorum", "Baskemolla", "Petersborg", "Kvesarum",
                 "Sodra_Rorum", "Bollerup", "Vagasked", "Klingstorp", "Tjurko",
                 "Vasternas", "Hasslo", "Aspo_Madviksvagen", "Sturko", "Gosslunda",
                 "Kleva_Strandskog", "Ynde", "Avaskar", "Hovmantorp", "Uras",
                 "Torne", "Brosarps_Backar", "Animskog_Cliff", "Animskog_Creek", "Aspo_Krokvagen",
                 "Billebjer", "Bokhultet", "Faringtofta", "Hjortahammar", "Jarnavik",
                 "Lyckas", "Mocklo", "Mollerod", "Sjuhalla", "Skaraskog",
                 "Svartbacksmala", "Taxas", "Varaskruv")

fgac_bypop_byyear <- expand_grid(Year = years, Population = populations) %>%
  pmap_dfr(function(Year, Population) { #alternate to lapply
    
    # Filter for specific year-population combo
    df_filtered <- fgac_allpop %>%
      filter(Year == !!Year, Population == !!Population)
    
    # Skip if too few observations to run specaccum
    if (nrow(df_filtered) < 2) return(NULL)
    
    mat <- df_filtered %>%
      group_by(Population_Bout, Pollinator_Group_Collapsed) %>%
      summarise(Total_Individuals = sum(Individuals, na.rm = TRUE), .groups = "drop") %>%
      pivot_wider(id_cols = Population_Bout,
                  names_from = Pollinator_Group_Collapsed,
                  values_from = Total_Individuals) %>%
      ungroup() %>%
      select(-Population_Bout) %>%
      replace(is.na(.), 0)
    
    # Skip if matrix has too few rows
    if (nrow(mat) < 2) return(NULL)
    
    fgac <- specaccum(mat, method = "random", permutations = 100)
    
    data.frame(
      Population = Population,
      Year       = as.factor(Year),
      bouts      = fgac$sites,
      richness   = fgac$richness,
      sd         = fgac$sd
    )
  })

```

### Plotting all Species Accumulation Curves (sep years)
```{r sac-bypop-byyear, message=FALSE, warning=FALSE}

ggplot(sac_bypop_byyear, aes(x = bouts, y = richness, color = Population, fill = Population)) +
#convert to hours for more meaningful plots
  #ggplot(sac_bypop_byyear, aes(x = (bouts*10)/60), y = richness, color = Population, fill = Population)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  facet_wrap(~ Year, scales = "free_x") +  # free_x because years may have different n bouts
#labs(x = "Observation Hours", y = "Species (N)",
  labs(x = "Observation Bouts", y = "Species (N)",
       title = "Rarefaction Curves by Population and Year") +
  theme_classic()+
  theme(legend.position = "right",
        legend.text = element_text(size = 5))

ggsave("rarefaction_bypop_sp.png", width = 20, height = 10, dpi = 300)

#FUNCTIONAL GROUPS
ggplot(fgac_bypop_byyear, aes(x = bouts, y = richness, color = Population, fill = Population)) +
# convert bouts to hours
#ggplot(fgac_bypop_byyear, aes(x = (bouts*10)/60, y = richness, color = Population, fill = Population)) +
  geom_ribbon(aes(ymin = richness - 2*sd, ymax = richness + 2*sd), 
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  facet_wrap(~ Year, scales = "free_x") +  # free_x because years may have different n bouts
#labs(x = "Observation Hours", y = "Species (N)",
  labs(x = "Observation Bouts", y = "Functional Groups (N)",
       title = "Rarefaction Curves by Population and Year") +
  theme_classic()
ggsave("rarefaction_bypop_fg.png", width = 20, height = 10, dpi = 300)
```
